Behavior Driven Development in Haskell

### 2012-05-25 paper.tex
s/annotation/markup/

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines annotation syntax in Haskell's comment.
Haddock provides code block annotations indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block annotation:

@@

To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use \hspec{} aslo integrated with QuickCheck.

@@

\begin{itemize}
\item Define a new Haddock annotation syntax for examples.
\item Define a new Haddock annotation syntax for properties.
\item Bring {\tt doctest} from the Python community to the Haskell
community and integrate it with QuickCheck.
\end{itemize}

First, we introduce a new Haddock annotation syntax for the example.
In this paper, we use {\tt >>>} which is the same as Python's one.
Here is an example:

@@

To specify properties in Haddock annotations, we also propose a new syntax. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:

@@

Examples and properties in Haddock annotations are automatically
tested by the {\tt doctest} command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

As we described, Haddock version 2.7.2 or earlier has
anotation syntax for code block but does not provide
annotation syntax for examples and properties.
We implemented annotation syntax for examples ({\tt >>>}) and 
it is available in Haddock version 2.8 or later.
We also extended Haddock to support annotation syntax for
properties ({\tt prop>})
which will be available in the future release.
Both examples and properties in Haddock annotation are pretty printed when producing document
such as in HTML.

@@

TBD(@sol): how to obtain examples and properties from Haddock annotations.
Old version uses Haddock API but now uses GHC API directly. Why?

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write visible behaviors in Haddock annotation:

@@

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
if Haskell compilers can infer correctly.
Also, we can cut and paste examples and their results
from Haskell interpreters to their Haddock annotations.

@@

We are grateful to David Waern, 
one of the Haddock maintainers, 
for reviewing two new syntax to Haddock.
We would like to thank
Nick Smallbone for having improved QuickCheck so that no-parameter properties are tested just once.

>>>>

The standard documentation mechanism for Haskell is
Haddock~\cite{haddock}.
Haddock defines markup syntax in Haskell's comment.
Haddock provides code block markup indicated with ({\tt >}).
For example, the {\tt container} packages has the following code block markup:

@@

To specify visible behavior, we propose to add two new markups to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use \hspec{} aslo integrated with QuickCheck.

@@

\begin{itemize}
\item Define a new Haddock markup for examples.
\item Define a new Haddock markup for properties.
\item Bring {\tt doctest} from the Python community to the Haskell
community and integrate it with QuickCheck.
\end{itemize}

First, we introduce a new Haddock markup for the example.
In this paper, we use {\tt >>>} which is the same as Python's one.
Here is an example:

@@

To specify properties, we also propose a new markup. In this paper, we use the {\tt prop>} prefix. We can write the no-parameter equation as follows:

@@

Examples and properties in Haddock markup are automatically
tested by the {\tt doctest} command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

As we described, Haddock version 2.7.2 or earlier has
the markup for code block but does not provide
markups for examples and properties.
We implemented a markup for examples ({\tt >>>}) and 
it is available in Haddock version 2.8 or later.
We also extended Haddock to support a markup for
properties ({\tt prop>})
which will be available in the future release.
Both examples and properties in Haddock markup are pretty printed when producing document
such as in HTML.

@@

TBD(@sol): how to obtain examples and properties from Haddock markup.
Old version uses Haddock API but now uses GHC API directly. Why?

@@

\noindent Note that the placeholders library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write visible behaviors in Haddock markup:

@@

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
if Haskell compilers can infer correctly.
Also, we can cut and paste examples and their results
from Haskell interpreters to their Haddock markup.

@@

We are grateful to David Waern, 
one of the Haddock maintainers, 
for reviewing two new markups to Haddock.
We would like to thank
Nick Smallbone for having improved QuickCheck so that no-parameter properties are tested just once.
